---
title: "Discussion 5"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
```

# Preliminaries

- This file should be in `STAT240/discussion/ds05`.
- You should also download `billboard_2000_weekly.csv` and `billboard_2000_dates.csv` to `STAT240/data`.

- This assignment should be completed in your **assigned Canvas groups**. Only one person has to submit on behalf of the group. You may submit whatever you have at the end of discussion section for full credit.

# Present Group Members

Please **write the names of the present group members** here.

> Rowan Tichelaar

> Vibha Nair

> Avery Dunham

> 

# Background

```{r}
billboard_weekly = read_csv("../../data/billboard_2000_weekly.csv")
```

`billboard_weekly` contains data for 317 tracks which appeared in the Billboard Top 100 (most popular songs in the United States) at any time in the year 2000.

After the song enters the Billboard Top 100, its `rank` in the top 100 is recorded (1 is the most popular song of all) for the next 76 `week`s. 

If a song is still in the Top 100 in that `week` since its debut in the Top 100, its `rank` is recorded (1 is the most popular song of all).

Many do not last in the top 100 for long; `NA` means that song was not in the Billboard Top 100 in that week of its lifespan. 

For example, in the first ten rows of `billboard_weekly`: 2 Pac's "Baby Don't Cry" entered the Top 100 at #87 in February 2000, and peaked at #72 three weeks later. Eight weeks after it entered, it dropped out of the Top 100 entirely.

```{r}
billboard_weekly
```

# Problem 1

Using a `pivot` command, **create and save** a new dataframe called `billboard_tracks`, starting from `billboard_weekly`.

`billboard_tracks` should have one row PER TRACK (for 317 tracks), and columns `artist`, `track`, `date.entered`, `wk1`, `wk2`, ... `wk76`, for a total of 79 columns. 
The columns `wk1`, `wk2`, ..., `wk76` should indicate that track's rank in that week.

```{r}
billboard_tracks <- billboard_weekly %>%
  pivot_wider(
    id_cols = c(artist, track, date.entered),
    names_from = week,
    values_from = rank
  )

print(billboard_tracks)
```

# Problem 2

All of the following problems purposefully do not specify whether you should use `billboard_weekly` or `billboard_tracks`, or how to get the answer. This is left up to you to figure out. None of the problems rely on each other from here on out

You should answer by starting with one of those dataframes, performing some manipulation, and letting the output print to console, no need to save any output from here on out.

---

**Which track had the best ranking (lowest number) in its very first week in the Billboard Top 100?**

```{r}
billboard_tracks %>%
  select(artist,track,wk1) %>%
  arrange(wk1)%>%
slice(1)
```

# Problem 3

Is there any `track` name that got used by two different artists?

```{r}
tracks_with_multiple_artists <- billboard_weekly %>%
  distinct(artist, track) %>%  # Unique artist-track combinations
  group_by(track) %>%
  filter(n() > 1) %>%
  ungroup()

tracks_with_multiple_artists

```

# Problem 4

Three songs that entered the Top 100 in 2000 stayed there all the way to April 7th, 2001.

```{r}
billboard_dates = read_csv("../../data/billboard_2000_dates.csv")

longest_running_songs = slice_max(billboard_dates, date.of.week, n = 3)

longest_running_songs
```

**What were their rankings in that final week?** Your answer should use a `*_join` command to join `longest_running_songs` to EITHER `billboard_weekly` or `billboard_tracks`.

```{r}

```

# Problem 5

**Which track had the biggest rise (e.g. their ranking "dropped" towards 1 the most) from week 1 to week 5?**

```{r}
billboard_tracks %>%
  mutate(rise=wk1-wk5) %>%
  arrange(desc(rise)) %>%
  select(artist,track,wk1,wk5,rise)%>%
  slice(1)
```


# Problem 6

**Which five tracks spent the most total weeks in the Billboard Top 100?**

```{r}
top_tracks_by_weeks <- billboard_tracks %>%
  mutate(total_weeks = rowSums(!is.na(select(., starts_with("wk"))))) %>%
  arrange(desc(total_weeks)) %>%
  select(artist, track, total_weeks) %>%
  slice(1:5)

top_tracks_by_weeks

```

# Problem 7

**Which tracks hit rank #1 at ANY point?**

```{r}
number_one_tracks <- billboard_weekly %>%
  filter(rank == 1) %>%
  distinct(artist, track)

number_one_tracks

```

# Problem 8

**Which track had the most weeks ranked at #1, and how many weeks was it there?**

```{r}
most_weeks_at_number_one <- billboard_weekly %>%
  filter(rank == 1) %>%
  group_by(artist, track) %>%
  summarise(weeks_at_number_one = n()) %>%
  arrange(desc(weeks_at_number_one)) %>%
  slice(1)

most_weeks_at_number_one

```

