---
title: "Discussion 7"
output: html_document
---

# Preliminaries

- This file should be in `STAT240/discussion/ds07` and the script `ggprob.R` should be in `STAT240/scripts` on your local computer.

- This assignment should be completed in your **assigned Canvas groups**. Only one person has to submit on behalf of the group. You may submit whatever you have at the end of discussion section for full credit.

# Present Group Members

Please **write the names of the present group members** here.

> Present Group Members: Jenny Kim, Avery Dunham, Vibha Nair

> Absent Group Members: Rowan Tichelaar 



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      error = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
library(tidyverse)
source("../../scripts/ggprob.R")
```



# Problem 1

For each scenario, explain why the random variable will or will not have a binomial distribution. If not, cite the specific BINS assumption(s) that are not met.

1. A bucket contains 10 colored tokens with five that are red and five that are blue. Four tokens are drawn at random from the bucket one at a time, but without replacing the tokens drawn. $X_1$ is the number of red tokens selected.

> The random variable  $X_1$  does not have a binomial distribution because the trials are not independent (tokens are drawn without replacement), and the probability of success (drawing a red token) is not constant.

2. A fair coin is tossed repeatedly until the tenth head is tossed. $X_2$ is the number of tails tossed prior to the tenth head.

> The random variable $X_2$ does not have a binomial distribution because the number of trials is not fixed. The process continues until the 10th head is tossed, which violates the fixed number of trials assumption.

3. Four buckets each contain a total of five tokens each, some red and some blue. The number of red tokens in the buckets are 1, 2, 3, and 4 with blue tokens making up the remainder. One token is drawn at random from each bucket. $X_3$ is the total number of red tokens drawn.

> The random variable $X_3$ does have a binomial distribution because there are fixed, independent trials, each with a constant probability of success (drawing a red token from each bucket). All BINS conditions are satisfied.

# Problem 2

For $X \sim \text{Binomial}(15,0.2)$ 

- Part a: Calculate the mean, variance and standard deviation of $X$.

```{r}
n <- 15
p <- 0.2

mean_X <- n * p

variance_X <- n * p * (1 - p)

std_dev_X <- sqrt(variance_X)

mean_X
variance_X
std_dev_X

```

- Part b: Calculate the probability of an outcome greater than or equal to 4. List all the ways you can find this probability.

```{r}
n <- 15
p <- 0.2

prob_X_greater_equal_4 <- 1 - pbinom(3, size = n, prob = p)

prob_X_greater_equal_4

```

# Problem 3

Let $X_1$ and $X_2$ be two draws from $X \sim N(0,1)$. Order the five events below based on which events are least to most likely to occur.

Event A: $X_1 < X_2$    
Event B: $X_1 = \mu$    
Event C: $\mu - \sigma < X_2 < \mu + \sigma$    
Event D: $X_1 > 1.5$ AND $X_2 > 1.5$

>EventB, EventD, EventA, EventC

# Problem 4

Create a data frame with columns `n`, `p`, `x`, `mu`, and `prob` where:

-   `n` varies from 2 to 100 by twos (so each `n` value is even);
-   `p` equals 0.5;
-   `x` is $n/2$;
-   `mu` is the mean of the distribution;
-   `prob` is $P(X = x)$

Show the first few rows using `head()`.

```{r}
n_values <- seq(2, 100, by = 2)  
p <- 0.5  
x_values <- n_values / 2  
mu_values <- n_values * p 

prob_values <- dbinom(x_values, size = n_values, prob = p)

df <- data.frame(n = n_values, p = p, x = x_values, mu = mu_values, prob = prob_values)

head(df)

```

**Part a** What is the relationship between `x` and `mu`?

>The relationship between x and mu is that x is exactly equal to mu. This is becuase x=n/2 and u=np, with p=0.5, so u=n/2, making x=u. 

**Part b** Make a line graph of `prob` versus `n`.

```{r}
library(ggplot2)

ggplot(df, aes(x = n, y = prob)) +
  geom_line() +
  labs(title = "Probability vs. n", x = "n", y = "Probability (P(X = x))")

```

**Part c** Describe the pattern: how does the probability that a random variable is equal to the mean change as $n$ increases?

> As n increases, the probability P(X=u) decreases because the distribution spreads out.
# Problem 5 

Find the probabilities of the area shaded in the graph below for $\text{Normal}(2.8, 6)$.

*Make sure you have run `source("../../scripts/ggprob.R")`  at the top of this file! If you get the error `cannot open file '../../scripts/ggprob.R': No such file or directory`, reset your working directory with `Session > Set Working Directory > To Source File Location`.*


1. Don't change the R chunk below

```{r}
gnorm(2.8, 6) +
  geom_norm_fill(2.8, 6, a = 1.5, b = 4, fill = "red") +
  geom_vline(xintercept = 1.5) +
  geom_vline(xintercept = 4) +
  annotate("text", x = -1.5, y = .07, label = "x = 1.5") +
  annotate("text", x = 6.5, y = .07, label = "x = 4")
```

```{r}
mu <- 2.8
sigma <- 6

prob_4 <- pnorm(4, mean = mu, sd = sigma)
prob_1.5 <- pnorm(1.5, mean = mu, sd = sigma)

prob_shaded <- prob_4 - prob_1.5

prob_shaded

```

2. Don't change the R chunk below

```{r}
gnorm(2.8, 6) +
  geom_norm_fill(2.8, 6, a = -2, b = 25, fill = "red") +
  geom_vline(xintercept = -2) +
  annotate("text", x = 1.5, y = .07, label = "x = -2") 
```

```{r}
mu <- 2.8
sigma <- 6
prob_25 <- pnorm(25, mean = mu, sd = sigma)
prob_neg_2 <- pnorm(-2, mean = mu, sd = sigma)

prob_shaded <- prob_25 - prob_neg_2

prob_shaded

```

3. Don't change the R chunk below

```{r}
mu = 2.8
sigma = 6
gnorm(2.8, 6) +
  geom_norm_fill(2.8, 6, a = mu-2*sigma, b =  mu+2*sigma, fill = "red") +
  geom_vline(xintercept =  mu-2*sigma) +
  geom_vline(xintercept =  mu+2*sigma) +
  annotate("text", x = 0.7, y = .07, label = "2 standard deviation") +
  scale_x_continuous(breaks = c(mu-3*sigma, mu-2*sigma, mu-sigma, mu, mu+sigma,mu+2*sigma, mu+3*sigma))
```

```{r}
mu <- 2.8
sigma <- 6

prob_upper <- pnorm(mu + 2 * sigma, mean = mu, sd = sigma)
prob_lower <- pnorm(mu - 2 * sigma, mean = mu, sd = sigma)

prob_within_2sd <- prob_upper - prob_lower

prob_within_2sd

```


# Problem 6

It is known that for Normal distributions, about 68% of the data will fall within one standard deviation, about 95% within two standard deviations, and about 99.7% within three standard deviations from the mean. 

Calculate these probabilities for $\text{Normal}(0,1)$ and display all three values.

```{r}
mu <- 0
sigma <- 1

prob_1sd <- pnorm(mu + sigma, mean = mu, sd = sigma) - pnorm(mu - sigma, mean = mu, sd = sigma)
prob_2sd <- pnorm(mu + 2 * sigma, mean = mu, sd = sigma) - pnorm(mu - 2 * sigma, mean = mu, sd = sigma)
prob_3sd <- pnorm(mu + 3 * sigma, mean = mu, sd = sigma) - pnorm(mu - 3 * sigma, mean = mu, sd = sigma)

prob_1sd
prob_2sd
prob_3sd

```






