---
title: "Discussion 12"
output: html_document
---

# Preliminaries

-   This file should be in `STAT240/discussion/ds12` on your local computer.

-   This assignment should be completed in your **assigned Canvas groups**. Only one person has to submit on behalf of the group. You may submit whatever you have at the end of discussion section for full credit.

-   This assignment is to be completed in the first 10-20 mins of the discussion session. The rest of the discussion time the students are to work on the group projects with their discussion groups.

# Present Group Members

Please **write the names of the present group members** here.

> Present Group Members: Avery Dunham, Vibha Nair, Rowan Tichelaar

> Absent Group Members: Jenny Kim

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      error = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
library(tidyverse)
```

## Problems

The questions in this discussion assignment use the boston marathon data set. The data set `bm_sample` contains a random sample of 100 men and 100 women, between the age of 35-39, who ran in and completed the 2010 Boston Marathon. Times are listed in minutes.

```{r}
# This code is provided for you
bm_sample = read_csv("../../data/bm_sample.csv")
spec(bm_sample)
```

### Problem 1

Create and display a table with a numerical summary of the data.
Calculate the sample size, mean time, standard deviation of the times, and standard error of the mean ($s/\sqrt{n}$) separately for each sex. This table will have one row for each sex.

```{r}
bm_summary <- bm_sample %>%
  group_by(Sex) %>%
  summarise(
    n = n(),
    mean_time = mean(Time),
    sd_time = sd(Time),
    se_mean = sd_time / sqrt(n)
  )

bm_summary
```

### Problem 2

Construct 95% confidence intervals for each sex seperately for the mean time to complete the 2010 Boston Marathon among all competitors of that sex aged 35-39.


```{r}
bm_conf_int <- bm_sample %>%
  group_by(Sex) %>%
  summarise(
    n = n(),
    mean_time = mean(Time),
    sd_time = sd(Time),
    se_mean = sd_time / sqrt(n),
    t_critical = qt(0.975, df = n - 1),
    margin_error = t_critical * se_mean,
    ci_lower = mean_time - margin_error,
    ci_upper = mean_time + margin_error
  )
bm_conf_int %>%
  select(Sex, mean_time, ci_lower, ci_upper)
```


### Problem 3

**Part (a)** Perform hypothesis test at 10% significance level for the difference of mean or the two - sample `t.test()` for the mean time to complete the 2010 Boston Marathon for male is less than that for female participants. Write down all the steps of the hypothesis test.

```{r}
alpha <- 0.10

t_test_result <- t.test(Time ~ Sex, data = bm_sample,
                        alternative = "less",
                        var.equal = FALSE)

t_statistic <- t_test_result$statistic
degrees_freedom <- t_test_result$parameter
p_value <- t_test_result$p.value

results <- list(
  t_statistic = t_statistic,
  degrees_of_freedom = degrees_freedom,
  p_value = p_value
)
results

if (p_value < alpha) {
  decision <- "Reject the null hypothesis"
  conclusion <- "There is sufficient evidence to conclude that the mean marathon time for males is less than that for females among participants aged 35-39."
} else {
  decision <- "Fail to reject the null hypothesis"
  conclusion <- "There is not sufficient evidence to conclude that the mean marathon time for males is less than that for females among participants aged 35-39."
}

list(
  decision = decision,
  conclusion = conclusion
)

```


**Part (b)** Can you support the conclusion made for the hypothesis test based on the confidence interval obtained in the Step 4 and 5?

> Yes, the conclusion from the hypothesis test is supported by the confidence intervals obtained in Problem 2. The 95% confidence intervals for males ([209.91, 218.85] minutes) and females ([231.33, 240.96] minutes) do not overlap, indicating a significant difference. Therefore, the confidence intervals support the conclusion that males have a lower mean marathon time than females among participants aged 35-39.





