---
title: "Homework 3"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, error = TRUE)
library(tidyverse)
```

# Preliminaries

- This file should be in `STAT240/homework/hw03` on your local computer.
- You should also download `exoplanets-clean-through-2022.csv` to `STAT240/data` on your local computer.

```{r read-data, echo = FALSE}
planets = read_csv("../../data/exoplanets-clean-through-2022.csv")
```

**Problems in this homework may ask you to "start from" `planets`; do not overwrite what is in `planets`! That is, you should NOT start a line with `planets = ...`, rather do, `somethingElse = planets %>% ...`.**

# Problem 1

The columns `mass` and `radius` are relative to Earth. A mass or radius of 2.0 refers to a planet twice the mass or radius of Earth. A mass or radius of 0.5 refers to a planet half the mass or radius of Earth.

Starting from `planets`,

+ **Keep** only the planets whose `mass` AND `radius` are smaller than those of Earth's.

+ **Get rid of** the columns `id` and `star`. 

+ **Sort** the rows increasing by `mass`; the smallest mass should be the first row, and the largest mass should be the last row.

+ **Save** this dataframe as some name other than `planets`, and then in a separate line **print** the whole resulting table, using `print()` with the `n = Inf` argument.

```{r}
# Write your code here!
filtered_planets = planets %>%
  filter(mass < 1 & radius < 1) %>%
  select(-id, -star) %>%
  arrange(mass)
print(filtered_planets, n = Inf)

```

# Problem 2

Starting from `planets`:

+ **Keep** only planets discovered by the `method` "Radial Velocity".

+ **Keep** only planets whose `mass` and `radius` are both known. (e.g. they are not missing.)

+ **Group** and **summarize** that data, such that:
    - You get **one row per year**.
    - **Create** the columns `n_discovered` and `minimum_mass`, which contain the number of planets discovered in that year, and the smallest mass among planets discovered in that year.

+ **Save** this dataframe with the name `mass_by_year`. Then, on a separate line, let `mass_by_year` **print** as output so we can see the first ten rows in your knitted file (no need to use `print` to show the whole thing, we only need ten rows.)

```{r}
# Write your code here!
mass_by_year = planets %>%
  filter(method == "Radial Velocity", !is.na(mass), !is.na(radius)) %>%
  group_by(year) %>%
  summarize(n_discovered = n(), 
            minimum_mass = min(mass), 
            .groups = "drop") 
head(mass_by_year, 10)

```

Your first two rows should look like this if you did it correctly. (Column order is arbitrary, doesn't matter if `n_discovered` and `minimum_mass` are switched around, just match the values.)

```
    year n_discovered minimum_mass
   <dbl>        <int>        <dbl>
 1  1999            1       232.  
 2  2001            1      1392.  
```

# Problem 3

Using `mass_by_year` from the previous problem, **create a scatter plot** of `minimum_mass` on the y axis versus `year` on the x axis.

Incorporate the following customizations:

+ Map the `size` of the points to the number of discoveries in that year - namely, `n_discovered`.

+ Apply the `log10` transformation to the y-axis.
    - This can be done by adding `scale_y_log10()` as a layer to your plot, which is a shortcut for `scale_y_continuous(trans = "log10")`.

+ Improve the **legend title** and the **y axis label** so they are not just the variable names.

```{r}
# Write your code here!

ggplot(data = mass_by_year, mapping = aes(x = year, y = minimum_mass, size = n_discovered)) +
  geom_point() + 
  scale_y_log10() + 
  labs(y = "minimum mass", size = "Number of Discoveries")
```

# Problem 4

Starting from `planets`, the original dataframe from the top of the file,

**Print out the planet name, mass, radius, and density of the top five most dense planets.**

To do so, you will have to calculate the density of each planet first, and then find the top five by density.

The density of a planet is its mass divided by its volume.

The volume of a planet is $\frac{4}{3} \times \pi \times \text{radius}^3$. (More easily viewed in the .html file.) Note that `pi` is a constant bult into R that you may use.

```{r}
# Write your code here!
most_dense_planets = planets %>% 
  mutate(density = mass/((4*pi*(radius^3))/3)) %>% 
  select(planet, mass, radius, density) %>% 
  arrange(desc(density))

most_dense_planets %>% 
  head(5)
```

*Note: This question requires you to understand the request and figure out which commands to chain together. Previous questions have indicated the step by step process, it is intentionally left out of this question and some future ones.*

# Problem 5

Every row of `planets` represents a single planet. Multiple planets may be orbiting the same `star`, as in our solar system.

Which star or stars have the most planets orbiting them in this dataset? How many planets are orbiting that star or stars?

To answer this question, start from `planets`, then **create and print a dataframe** with columns `star` and `n`, with `n` representing how many planets are orbiting that `star`.

```{r}
# Write your code here!

star_planet_counts = planets %>%
  group_by(star) %>%
  summarize(n = n(), .groups = 'drop') %>%
  arrange(desc(n))

print(star_planet_counts)

```

# Problem 6

Problems 6 - 8 take you through a relatively complex analysis -> visualization process, which mimics what you might provide to a client asking the question:

"How has the most popular method of planet discovery changed over time?" 

---

Starting from `planets`,

+ **Keep** only the planets discovered by the `method`s "Imaging", "Microlensing", "Radial Velocity", or "Transit".

+ **Keep** only the planets discovered in 2000 or later.

+ **Count** how many times each remaining `method` was used within each `year`.
    - This will require you to group by both `method` and `year` at the same time.

+ **Sort** the rows such that the earliest year is first, and the most recent year is last.

+ **Save** this dataframe with the name `methods_within_year`, and then let `methods_within_year` be **printed as output** so the first ten rows are visible in your .html file.

```{r}
# Write your code here!

methods_within_year = planets %>%
  filter(method %in% c("Imaging", "Microlensing", "Radial Velocity", "Transit") & year >= 2000) %>%
  group_by(year, method) %>%
  summarize(n = n(), .groups = 'drop') %>%
  arrange(year, method)

methods_within_year <- as_tibble(methods_within_year)

print(methods_within_year, n = 10) 

```

The top of `methods_within_year` should look similar to this (column order does not matter if `method` or `year` or `n` are switched).

```
# A tibble: 78 × 3
    year method              n
   <dbl> <chr>           <int>
 1  2000 Radial Velocity    16
 2  2001 Radial Velocity    12
 3  2002 Radial Velocity    28
 4  2002 Transit             1
```

# Problem 7

Starting from `methods_within_year` from problem 6 above,

+ **Grouping by year, add a column** called `yearTotal`. `yearTotal` should indicate how many planets were discovered within that year across all methods.
    - Here's the first four rows to check your work against:
```
    year method              n yearTotal
   <dbl> <chr>           <int>     <int>
 1  2000 Radial Velocity    16        16
 2  2001 Radial Velocity    12        12
 3  2002 Radial Velocity    28        29
 4  2002 Transit             1        29
```

+ Now, **add another column** called `methodProportion`, which determines what percentage of the discoveries within that year were by that method.
    - Here's what `methodProportion` should look like, again, column order doesn't matter:
```
    year method              n yearTotal methodProportion
   <dbl> <chr>           <int>     <int>            <dbl>
 1  2000 Radial Velocity    16        16           1     
 2  2001 Radial Velocity    12        12           1     
 3  2002 Radial Velocity    28        29           0.966 
 4  2002 Transit             1        29           0.0345
```

**Save** this dataframe with the name `methods_within_year_proportions`, and then let `methods_within_year_proportions` be **printed as output** so the first ten rows are visible in your .html file.

```{r}
# Write your code here!
yearTotal = methods_within_year %>% 
  group_by(year) %>% 
  summarize(yearTotal = sum(n))

methods_within_year_proportions = methods_within_year %>% 
  left_join(yearTotal, by = "year") %>% 
  mutate(methodProportion = n/yearTotal)

methods_within_year_proportions %>% 
  head(10)

```


## Problem 8

Using `methods_within_year_proportions` from the previous question,

**Create a column plot** with `year` on the x axis and `methodProportion` on the y axis.

The height of each bar should be exactly 1. Now, use the `fill` aesthetic to color each column by its different `method`s.

```{r}
# Write your code here!

ggplot(data = methods_within_year_proportions, aes(x = year, y = methodProportion, fill = method)) + 
  geom_col() +
  labs(y = "Proportion of Discoveries", x = "Year", title = "Proportion of Discoveries by Method over Time")
```

Now, let's answer the client's question based on the graph. **"How has the most popular method of planet discovery changed over time?"**

> Early 2000s:Radial Velocity was the most popular method for discovering planets

>Mid-2000's to 2010s: There was a shift from Radial Velocity to the Transit method as technology improved

>Recent Year: The Transit method has become the most commonly used method for discovering new planets.
